Approximation as Differentiation

### 2025-03-31 notes/sec/biproduct.tex
copair-ε.
Construct biproducts from coproducts on a cmon-category.
Merge propositions 3.6 and 3.7.

\begin{proposition}
\item
\begin{itemize}
\item If $C$ has biproducts then it has products given by $(A \biprod B, \biproj_A, \biproj_B)$ and coproducts
given by $(A \biprod B, \biinj_A, \biinj_B)$.
\item If $C$ has products (or coproducts) then they are biproducts.
\item If $C$ has a terminal (or initial) object then it is a zero object.
\end{itemize}
\end{proposition}

\noindent We say that $C$ is \emph{semi-additive} if it has finite biproducts. Having products or coproducts
is sufficient for semi-additivity:

\begin{proposition}
Suppose $C$ is $\CMon$-enriched.
\begin{itemize}
\item If $C$ has products given by $(A \times B, \proj_1, \proj_2)$ then:
\begin{enumerate}
\item $\prodM{0_{A,B}}{0_{A,C}} = 0_{A,B \times C}$
\item $\prodM{f_1 + g_1}{f_2 + g_2} = \prodM{f_1 + f_2}{g_1 + g_2}$
\item $(A \times B, \biinj_A, \biinj_B, \proj_1, \proj_2)$ is a
biproduct with $\biinj_A: A \to A \times B = \prodM{\id_A}{0_{A,B}}$ and $\biinj_B: B \to A \times B =
\prodM{0_{A,B}}{\id_B}$.
\end{enumerate}
\item If $C$ has coproducts given by $(A + B, \inj_1, \inj_2)$ then
\begin{enumerate}
\item $(A + B, \inj_1, \inj_2, \biproj_A, \biproj_B)$ is a
biproduct with $\biproj_A: A + B \to A = \coprodM{\id_A}{0_{B,A}}$ and $\biproj_B: A + B \to B =
\coprodM{0_{A,B}}{\id_B}$.
\end{enumerate}
\end{itemize}
\end{proposition}

>>>>

\noindent We say that $C$ is \emph{semi-additive} if it has finite biproducts. Having products or coproducts
is sufficient for semi-additivity.

\begin{proposition}
\label{prop:biproduct:from-product-or-coproduct}
Suppose $C$ is $\CMon$-enriched.
\begin{itemize}
\item If $C$ has biproducts given by $(A \biprod B, \biinj_A, \biinj_B, \biproj_A, \biproj_B)$ then $(A
\biprod B, \biproj_A, \biproj_B)$ is a product and $(A \biprod B, \biinj_A, \biinj_B)$ is a coproduct.
\item If $C$ has binary products given by $(A \times B, \proj_1, \proj_2)$ then:
\begin{enumerate}
\item $\prodM{0_{A,B}}{0_{A,C}} = 0_{A,B \times C}$
\item $\prodM{f_1 + g_1}{f_2 + g_2} = \prodM{f_1 + f_2}{g_1 + g_2}$
\item $(A \times B, \biinj_A, \biinj_B, \proj_1, \proj_2)$ is a
biproduct with $\biinj_A: A \to A \times B = \prodM{\id_A}{0_{A,B}}$ and $\biinj_B: B \to A \times B =
\prodM{0_{A,B}}{\id_B}$.
\end{enumerate}
\item If $C$ has binary coproducts given by $(A + B, \inj_1, \inj_2)$ then
\begin{enumerate}
\item $\coprodM{0_{A,C}}{0_{B,C}} = 0_{A + B,C}$
\item $\coprodM{f_1 + g_1}{f_2 + g_2} = \coprodM{f_1 + f_2}{g_1 + g_2}$
\item $(A + B, \inj_1, \inj_2, \biproj_A, \biproj_B)$ is a
biproduct with $\biproj_A: A + B \to A = \coprodM{\id_A}{0_{B,A}}$ and $\biproj_B: A + B \to B =
\coprodM{0_{A,B}}{\id_B}$.
\end{enumerate}
\item If $C$ has a terminal (or initial) object then it is a zero object.
\end{itemize}
\end{proposition}
